---
title: "ES 207 Homework 7"
author: "Christiana Ade"
date: "April 22, 2018"
output: html_notebook
---

```{r}
# read in packages
require(tidyverse)
require(moments)
require(smwrBase)
require(lubridate)
library(dplyr)
require(corrplot)
require(stringr)
require(broom)

# personal resources
  # https://stackoverflow.com/questions/9723208/aggregate-summarize-multiple-variables-per-group-e-g-sum-mean?noredirect=1&lq=1
# https://stackoverflow.com/questions/26665319/removing-na-in-dplyr-pipe 
# https://stackoverflow.com/questions/45551918/how-to-remove-na-in-summarize-all-to-summarize-mutliple-col-at-the-same-time

```

### Question 1
**For the Water Years 2005-2012, develop a predictive model of the mean monthly Chl-a concentration in the California Bay Delta using  other mean monthly water quality variables.**

```{r}
# read in dataset 
wq <- read_csv("./Data/BayDeltaWQ.csv", col_names = TRUE, na = c("NA", "n/p", "n/a"), guess_max = 30000 )

## Tidy the dataset
# compute water year
# remove years that are not 2005 - 2012
# compute mean based on monthly for all columns 
# drop rows with NA's in Chl-a and any columns that just have NA values

wq <- wq %>% mutate(SampleDate = mdy(SampleDate)) %>%
  mutate(water_year = waterYear(SampleDate)) %>%
  mutate(month_year = format(SampleDate, "%Y-%m"))%>% 
  filter( water_year >= 2005 & water_year <= 2012 ) %>%
  group_by(StationCode, month_year) %>%
  summarise_if(is.numeric, mean, na.rm = T) %>% 
  drop_na(`Chlorophyll a`) %>% discard(~all(is.na(.x))) %>%
  map_df(~.x)

# # curious about how many NA's there are in each column
#  wq %>% summarise_all(funs(sum(is.na(.))))
# how many complete cases
 wq.cc <-  wq[complete.cases(wq), ]

               
```

```{r}
## Plot the remaining variables against chl-a
# faceted plots
wq %>%
  gather(-`Chlorophyll a`, key = "var", value = "value") %>%
  ggplot(aes(x = `Chlorophyll a`, y = value)) +
    geom_point() +
    facet_wrap(~ var, scales = "free") +
    theme_bw()

## Individual plots
myVarsList <- names(wq)
myVarsList <- myList[!myList %in% c("StationCode","Chlorophyll a","SiteDepth","Depth")]

myPlot <- function(x){
  ggplot(wq, aes(x = `Chlorophyll a`, y= wq[,x])) + geom_point() + scale_y_continuous(name = x) +
    scale_x_continuous(lim = c(0,50))
}

lapply(myVarsList, myPlot)

```
Not looking extremely promising for most of the variables. Several look as though you would want to use quantile regression, there are a few weak linear relationships. 


```{r}

# multiple correlations
wq2 <- dplyr::select(wq, `Chlorophyll a`, myVarsList)

# Check covariance 
cov(wq2,use = "pairwise.complete.obs")

# check correlation 
pearCor <- as.data.frame(cor(wq2, use = "pairwise.complete.obs"))
kCor <- as.data.frame(cor(wq2, use = "pairwise.complete.obs", method = "kendall"))
sCor   <- as.data.frame(cor(wq2, use = "pairwise.complete.obs", method = "spearman"))

# linear model all data
lm.all <- lm(`Chlorophyll a` ~ ., data = wq2)
 summary(lm.all)
# step all data
 lms <- step(lm(`Chlorophyll a` ~ ., data = wq2))
lms
```

While I could fit a model with all 25 of these variables. I find it unlikely given the correlation tests above and the scatter plots that all of them would contribute to a "good" model. Therefore, I am going to reduce the variables that I test a step model on using the above correlations.

```{r}
# Intial variable reduction 
r <-  subset(pearCor,`Chlorophyll a` >= 0.25 )
x <- subset(kCor,`Chlorophyll a` >= 0.25 )[,1]
s <- subset(sCor,`Chlorophyll a` >= 0.25 )[,1]

r
x
s

```

Based on the above information- Fluorescence, pH, Kjeldahl Nitrogen (Total), Pheophytin a, Temperature will be included in my step model. I am aware that this is not the most accurate to go about this. For instance, some of the variables most likely do not have linear relationships and thus these correlation statistics are not applicable. Also, the cut off point of 0.25 is pretty arbitrary, but was used because of the low correlation values accross the board.

I am honestly really struggling with this. In order to do a step model, I would think that I need complete cases so that the observations all line up appropriately. 
Either way from my modeling selection if I use wq.cc (complete cases) or all the data then the best model I am getting is with all 5 variables included. However, individual models were not included with temperature and ph etc which would be problematic. I must really be messing something up.

```{r}
# individual models
lm.null <- lm(`Chlorophyll a` ~ 1, data = wq.cc); summary(lm.null)
lm.pa <- lm(`Chlorophyll a` ~ `Pheophytin a`, data = wq.cc); summary(lm.pa)
lm.pa.ph <- lm(`Chlorophyll a` ~ `Pheophytin a` + pH, data = wq2); summary(lm.pa.ph)
lm.pa.ph.fl <- lm(`Chlorophyll a` ~ `Pheophytin a` + pH + Fluorescence, data = wq.cc); summary(lm.pa.ph.fl)
lm.pa.ph.fl.t <- lm(`Chlorophyll a` ~ `Pheophytin a` + pH + Fluorescence + Temperature, data = wq.cc); summary(lm.pa.ph.fl.t)
lm.pa.ph.fl.t.n <- lm(`Chlorophyll a` ~ `Pheophytin a` + pH + Fluorescence + Temperature + `Kjeldahl Nitrogen (Total)`, data = wq.cc); summary(lm.pa.ph.fl.t.n)

# broom 
lms <- list(null=lm.null, chl.pa = lm.pa, chl.pa.ph = lm.pa.ph, chl.pa.ph.fl = lm.pa.ph.fl, chl.pa.ph.fl.t = lm.pa.ph.fl.t, chl.pa.ph.fl.t.n = lm.pa.ph.fl.t.n)
lms.stats <- mapply(glance, lms)
colnames(lms.stats) <- names(lms)
lms.stats

# lowest AIC and BIC are apparently when all the variables are included for complete cases, which is only two stations




```



```{r}

# step wise model for complete cases

modStep <- step(lm(`Chlorophyll a` ~ Fluorescence + pH + `Kjeldahl Nitrogen (Total)` + `Pheophytin a` + Temperature , data = wq3))

plot(wq.cc$`Chlorophyll a`,wq.cc$pH)
plot(wq.cc$`Chlorophyll a`,wq.cc$`Pheophytin a`)
plot(wq.cc$`Chlorophyll a`,wq.cc$Temperature)


#anova to test variance of variables
anova(lm.pa.ph.fl.t.n)

```

```{r}
# related variables
pairs(~`Chlorophyll a` + Fluorescence + pH + `Kjeldahl Nitrogen (Total)` + `Pheophytin a` + Temperature, data = wq.cc)
```

Alright, I am assuming that I have really messed something up. I would think that of the variables I picked that Pheophytin a is the best predictor for Chlorophyll-a because they are similar. if you look at the AIC in the step though Temperautre and pH also seem to be important and it is likely that we can remove Nitrogen as a variable. If you wanted to better account for variables that are related then you would want to compute a VIF test to test for multi-collinearity. If we look at the anova for the linear model with all 5 variables (lm.pa.ph.fl.t.n) (really bad name I know), then we can determine that most of the variance is explained by Pheophytin a from looking at the F value. 



### Part 2

I am going to assume that my best variable was Pheophytin a

```{r}
# label wet and dry years
# april to oct is dry 
# nov - feb is wet
wq.p2 <- wq3 %>% mutate(month = as.numeric(substr(month_year,6,7))) %>%
  mutate(season = ifelse(month >= 4 & month <= 9,"dry","wet" ) )
                         
``` 

```{r}
# model separately

chl.m1 <- lm(`Chlorophyll a`~ season, data = wq.p2); summary(chl.m1)
chl.m2 <- lm(`Chlorophyll a` ~ `Pheophytin a`, data = wq.p2); summary(chl.m2)
chl.m3 <- lm(`Chlorophyll a` ~ `Pheophytin a` + season, data = wq.p2); summary(chl.m3)

# most variance 
anova(chl.m3)

# AIC
AIC(chl.m1)
AIC(chl.m2)
AIC(chl.m3)
```
Much less of the variance is explained by the season than Pheophytin a. The best model is both together. 
I tested it out with Temperature and those models are much more closely related


```{r}
# differences between means of the seasons
wq.means <- wq.p2 %>% group_by(season) %>% summarise_if(is.numeric, mean, na.rm = T) 
wq.means$`Chlorophyll a`

# t- test on wet vs dry
wq.wet <- filter(wq.p2, season == "wet")
wq.dry <- filter(wq.p2, season == "dry")

# signficantly different
t.test(wq.wet$`Chlorophyll a`,wq.dry$`Chlorophyll a`)

# but are they normally distributed
wilcox.test(wq.wet$`Chlorophyll a`,wq.dry$`Chlorophyll a`)
```



The residuals are not normally distributed
```{r}
# residuals calculation
chl.m3.res <-  rstandard(chl.m3) # can also use function residual()
qqnorm(chl.m3.res)
qqline(chl.m3.res)
```


```{r}

```


